---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# tiedr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.com/yutannihilation/tiedr.svg?branch=master)](https://travis-ci.com/yutannihilation/tiedr)
<!-- badges: end -->

tiedr is an experimental package to "pack" (or "bundle") some columns to a data.frame column.

The idea behind this package is basically the same as `pack()`/`unpack()`, which was proposed on the tidyr's issue below. I didn't notice this issue at the time of deciding the main function name as `bundle()`...

* <https://github.com/tidyverse/tidyr/issues/523>

If you wonder how this is useful, my blog post might help:

* <https://yutani.rbind.io/post/enhancing-gather-and-spread-by-using-bundled-data-frames/>


## Usages

### `bundle()` and `unbundle()`

`bundle()` packs columns into a data.frame column, and `unbundle()` unpacks it.

```{r}
library(tiedr)

iris <- tibble::as_tibble(iris)

# bundle() takes unnamed arguments
iris %>% 
  bundle(-Species)

# bundle() also takes named arguments
d <- iris %>% 
  bundle(Sepal = starts_with("Sepal"),
         Petal = starts_with("Petal"))

# bundled columns can be bundled
d %>%
  bundle(-Species)

# unbundle()
unbundle(d, Sepal, Petal, sep = NULL)
```

### `auto_bundle()`

```{r}
d <- tibble::tribble(
  ~A_a_x, ~A_a_y, ~A_b_x, ~B_a_x, ~B_b_x, ~B_b_y,
       1,      2,      3,      4,      5,      6,
       2,      3,      4,      5,      6,      7
)

auto_bundle(d, everything())
auto_bundle(d, starts_with("A"))
```

### `gather_bundles()`

```{r}
auto_bundle(d, everything()) %>%
  gather_bundles()

auto_bundle(d, everything()) %>%
  gather_bundles(.key = "key_outer") %>%
  gather_bundles(.key = "key_inner")

auto_bundle(d, everything()) %>%
  gather_bundles(.key = "key_outer") %>%
  gather_bundles(.key = "key_inner") %>%
  gather_bundles(x:y, .key = "xy")
```

### `spread_bundles()`

```{r}
# https://github.com/tidyverse/tidyr/issues/149
d <- tibble::tribble(
  ~hw,   ~name,  ~mark,   ~pr,
  "hw1", "anna",    95,  "ok",
  "hw1", "alan",    90, "meh",
  "hw1", "carl",    85,  "ok",
  "hw2", "alan",    70, "meh",
  "hw2", "carl",    80,  "ok"
)

spread_bundles(d, name, c(mark, pr))

spread_bundles(d, name, c(mark, pr)) %>%
  unbundle(-hw)

# without prefix
spread_bundles(d, name, c(mark, pr)) %>%
  unbundle(-hw, sep = NULL)


# ?spread
tibble::tibble(x = c("a", "b"), y = c(3, 4), z = c(5, 6)) %>%
  spread_bundles(x, y)
```
